\( \tool \): JavaScript IR-based Semantics Extraction Toolchain

### 2020-04-13 ase20/compiler.tex
[ASE] Merged Section 4 & 5

\subsection{Token List Parser}
The token list parser is defined with \textit{parsing rules} written in extended
Scala parser combinators. We modified the meaning of the alternative composition
operator ( \( | \) ) to collect all the longest matched results. If a parser
detects a step that cannot be parsed or parsed in multiple ways, it reports the
step with the parsing results.  We use Packrat parsing~\cite{packrat} for
parsing rules, which support left recursion.

We provide two kinds of basic parsing rules: \textit{tag-based rules} and
\textit{content-based rules}.  A tag-based rule just checks whether the next
token has a given tag.  For example, the tag-based parser \( \code{varT} \) and
\( \code{codeT} \) check whether the next token has the tag \( \code{<var>} \)
and \( \code{<code>} \), respectively.  A content-based parser checks whether
the next token is a text token and its content passes a given condition.  For
example, the string literal \( \code{"Perform"} \) denotes a content-based
parser that checks whether the next token is a text token with the content \(
\code{Perform} \).  We also provide two content-based parsers \( \code{word} \)
and \( \code{number} \) that check whether the content of the next token
consists of only alphabets or numbers, respectively.  In parsing rules, all the
helper functions defined in Scala parser combinators are available.  For
instance, the helper function \( \code{repsep(p, q)} \) generates a new parsing
rule that denotes zero or more repetition of the parsing rule \( \code{p} \)
using another parsing rule \( \code{q} \) as a separator.

Consider the following example parsing rule for the step 5 of the
\textbf{Evaluation} algorithm in Figure~\ref{fig:array-literal-eval}(a).
\begin{lstlisting}[style=myScalastyle]
// statements
val Stmt = "Perform" ~ Expr ~ "." ^^ ...
// expressions
val Expr =
  // codes        // false literal
  codeT    ^^ ... |  "false"           ^^ ... |
  // variables       // additions
  varT     ^^ ... |  Expr ~ "+" ~ Expr ^^ ... |
  // function calls
  word ~ "(" ~ repsep(Expr, ",") ~ ")" ^^ ...
\end{lstlisting}
We omit the conversion rule for each compile rule for the brevity.  The \(
\code{Stmt} \) compile rule  describes how to compile statements with a single
parsing rule, and the \( \code{Expr} \) compile rule describes how to compile
expressions with five parsing rules.  A token parser with the above rules parses
the step 5 of \textbf{Evaluation} to the following token AST:
\begin{center}
  \includegraphics[width=0.5\textwidth]{img/token-ast.pdf}
\end{center}

@@

\item \textbf{Abstract data types:} \( \ires \) supports only three abstract
data types: \( \code{Record} \) for mappings from some values to other values,
\( \code{List} \) for sequential data, and \( \code{Symbol} \) for singleton
data.  For example, ECMAScript environment records are represented as \(
\code{Record} \) from string values into addresses that represent other types
of \( \code{Record} \) for bindings.
\end{itemize}
We define syntax of \( \ires \) consisting 15 kinds of instructions and 26 kinds
of expressions with the notation \( \irinst \) and \( \irexpr \), respectively.
Moreover, we also formally define their operational semantics with the form \(
\irstate \vdash \irinst \Rightarrow \irstate \) for instructions, and the form
\( \irstate \vdash \irexpr \Rightarrow (\irvalue, \irstate) \) for expressions
where \( \irvalue \) denotes a value and \( \irstate \) denotes a state.
However, we omit this formalization in this paper for brevity, instead, we
include it in the companion material\footnote{\inred{???}}.

>>>>

\subsection{Token List Parser}
The token list parser is defined with \textit{parsing rules}. Each parsing rule
is a basic parsing rule or a composition of other parsing rules. The composition
\code{A | B} of two parsing rules \code{A} and \code{B} parses the input using
both rules and collect the longest matched results.  If both rules are failed or
matched the same length of the input, the composition is failed.

We provide two kinds of basic parsing rules: \textit{tag-based rules} and
\textit{content-based rules}.  A tag-based rule just checks whether the next
token has a given tag.  For example, the tag-based parser \( \code{varT} \) and
\( \code{codeT} \) check whether the next token has the tag \( \code{<var>} \)
and \( \code{<code>} \), respectively.  A content-based parser checks whether
the next token is a text token and its content passes a given condition.  For
example, the string literal \( \code{"Perform"} \) denotes a content-based
parser that checks whether the next token is a text token with the content \(
\code{Perform} \).  We also define two content-based parsers \( \code{word} \)
and \( \code{number} \) that check whether the content of the next token
consists of only alphabets or numbers, respectively.  In addition, we provide
several helper functions in parsing rules such as optional rule \code{A?},
positive(negative) predicate \code{+A}(\code{-A}).  For instance, the helper
function \( \code{repsep(A, B)} \) generates a new parsing rule that denotes
zero or more repetition of the parsing rule \( \code{A} \) using another parsing
rule \( \code{B} \) as a separator.

Consider the following example parsing rule for the step 5 of the
\textbf{Evaluation} algorithm in Figure~\ref{fig:array-literal-eval}(a).
\begin{lstlisting}[style=myScalastyle]
// statements
val Stmt = "Perform" ~ Expr ~ "." ^^ ...
// expressions
val Expr =
  // codes        // false literal
  codeT    ^^ ... |  "false"           ^^ ... |
  // variables       // additions
  varT     ^^ ... |  Expr ~ "+" ~ Expr ^^ ... |
  // function calls
  word ~ "(" ~ repsep(Expr, ",") ~ ")" ^^ ...
\end{lstlisting}
We omit the conversion rule for each compile rule for brevity.  The \(
\code{Stmt} \) compile rule  describes how to compile statements with a single
parsing rule, and the \( \code{Expr} \) compile rule describes how to compile
expressions with five parsing rules.  A token parser with the above rules parses
the step 5 of \textbf{Evaluation} to the following token AST:
\begin{center}
  \includegraphics[width=0.5\textwidth]{img/token-ast.pdf}
\end{center}

@@

\item \textbf{Abstract data types:} \( \ires \) supports only three abstract
data types: \( \code{Record} \) for mappings from some values to other values,
\( \code{List} \) for sequential data, and \( \code{Symbol} \) for singleton
data.  For example, ECMAScript environment records are represented as \(
\code{Record} \) from string values into addresses that represent other types
of \( \code{Record} \) for bindings.
\end{itemize}
We define syntax of \( \ires \) including 15 kinds of instructions and 26 kinds
of expressions with the notation \( \irinst \) and \( \irexpr \), respectively.
Moreover, we also formally define their operational semantics with the form \(
\irstate \vdash \irinst \Rightarrow \irstate \) for instructions, and the form
\( \irstate \vdash \irexpr \Rightarrow (\irvalue, \irstate) \) for expressions
where \( \irvalue \) denotes a value and \( \irstate \) denotes a state.
However, we omit this formalization in this paper for brevity, instead, we
include it in the companion material\footnote{\inred{???}}.

\subsection{Implementation}
We implemented {\sf Algorithm Compiler} by extending Packrat
parsing~\cite{packrat} library in Scala parser combinators. We modified the
meaning of the composition operator ( \( | \) ) to collect all the
longest matched results.  If a parser detects a step that cannot be parsed or
parsed in multiple ways, it reports the step with the parsing results.

\paragraph{Compile Rules}
\textsf{Algorithm Compiler} requires compile rules to compile given abstract
algorithms to \( \ires \) functions.  As already explained in
Section~\ref{sec:overview}, we found general patterns in writing style of
abstract algorithms.  We manually define the general compile rules to represent
such writing style with six different kinds as summarized in
Table~\ref{table:rules}.  The compile rule for statements, \( \code{Stmt} \),
generates \( \ires \) instructions.  The \( \code{Expr} \), \( \code{Cond} \),
and \( \code{Value} \) compile rules generate \( \ires \) expressions, but they
represents different contexts in ECMAScript; \( \code{Expr} \) represents a
context where any expression can appear, \( \code{Cond} \) denotes a context
where any boolean-valued expression can appear, and \( \code{Value} \)
represents a context where a fully evaluated value can appear.  The \( \code{Ty}
\) compile rule denotes type names and generates string primitives used in
object constructions.  The \( \code{Ref} \) represents references such as
identifier lookup and member accesses of objects, and it generates \( \ires \)
references.

\begin{table}[t]
  \centering
  \caption{General compile rules for ECMAScript}
  \label{table:rules}
\vspace*{-1em}
  \[
  \begin{array}{c?r|r|r|r|r|r}%|r}
      \text{Name}
      & \multicolumn{1}{c|}{\code{Stmt}}
      & \multicolumn{1}{c|}{\code{Expr}}
      & \multicolumn{1}{c|}{\code{Cond}}
      & \multicolumn{1}{c|}{\code{Value}}
      & \multicolumn{1}{c|}{\code{Ty}}
      & \multicolumn{1}{c}{\code{Ref}}\\\hline
      \text{\# Rules}
      & \inred{17}
      & \inred{16}
      & \inred{8}
      & \inred{11}
      & \inred{33}
      & \inred{7}
    \end{array}
  \]
\vspace*{-2em}
\end{table}

\paragraph{Global Setting}
To construct {\sf AST-\( \ires \) Translator}, global settings are also required
and it consists of \textit{ECMAScript data types} and \textit{built-in objects}.
Unlike compile rules, global settings are slightly different depending on
ECMAScript versions. In this paper, we construct global settings only for the
most recent ECMAScript version, ES10.

ECMAScript describes data types with several fields and methods used in the
specification.  Their methods are abstract algorithms but their semantics are
slightly different with other. They implicitly get their receiver objects as
arguments at callsites.  To mimic such implicit behavior, we added a special
variable \( \code{this} \) as the first parameter of a method, and passed a
receiver object at its callsite by modifying \textsf{Algorithm Compiler}.  For
example, an Environment Record type has the \textsf{DeleteBinding (N)} method.
Then, it gets two parameters, the special parameter \( \code{this} \) and normal
one \( \code{N} \).  Thus, the method call
\raisebox{-0.65ex}{\includegraphics[width=0.17\textwidth]{img/method-call.png}}
is compiled into the \( \ires \) instruction: \code{(DclRec.DeleteBinding
DclRec, N)}

In ECMAScript, built-in objects are pre-defined functions with several
built-in functions.  For example, \( \code{Array} \)
is the constructor of array objects, and its prototype
\( \code{Array.prototype} \) has built-in functions for array objects.
For instance, \( \code{[1,2,3].flat()} \) calls the
\( \code{Array.prototype.flat} \) built-in function with the array
\( \code{[1,2,3]} \).  Because built-in functions are also abstract
algorithms, each of them is automatically converted to an \( \ires \)
function.  However, the structures of built-in objects should be
manually implemented.  Thus, we implemented built-in objects in Scala
and connected their properties with the extracted \( \ires \) functions.
%
Some built-in objects that are explicitly referenced in abstract
algorithms are intrinsic objects, which have their own
aliased names summarized in Table 7 of Section 6.1.7.4
\textsf{Well-Known Intrinsic Objects}.  We extracted the alias
into \textsf{Global Setting} to utilize it during evaluation.

### 2020-04-13 ase20/overview.tex
[ASE] Merged Section 4 & 5

\section{Overview}
\begin{figure}
  \centering
  \includegraphics[width=0.48\textwidth]{img/overview.pdf}
  \caption{Overall structure of \( \tool \): Automatically synthesized {\sf
    JavaScript Parser} and {\sf AST-\( \ires \) Translator} for JavaScript
    IR-based semantics}
  \label{fig:overview}
\vspace*{-1em}
\end{figure}

@@

In the remainder of this paper, we explain the details of how to
automatically generate parsers (Section~\ref{sec:parser}) and how to compile
abstract algorithms (Section~\ref{sec:compiler}).  After describing implementation
details (Section~\ref{sec:impl}), we evaluate \( \tool \)
(Section~\ref{sec:eval}), discuss related work (Section~\ref{sec:related}), and
conclude (Section~\ref{sec:conclude}).

>>>>

\section{Overview}\label{sec:overview}
\begin{figure}
  \centering
  \includegraphics[width=0.48\textwidth]{img/overview.pdf}
  \caption{Overall structure of \( \tool \): Automatically synthesized {\sf
    JavaScript Parser} and {\sf AST-\( \ires \) Translator} for JavaScript
    IR-based semantics}
  \label{fig:overview}
\vspace*{-1em}
\end{figure}

@@

In the remainder of this paper, we explain the details of how to
automatically generate parsers (Section~\ref{sec:parser}) and how to compile
abstract algorithms (Section~\ref{sec:compiler}).  After evaluating \( \tool \)
(Section~\ref{sec:eval}), we discuss related work (Section~\ref{sec:related}), and
conclude (Section~\ref{sec:conclude}).
